Changing Flights In The Air\\ (Abortable Chain Agreement)

### 2015-05-14 paper/intro.tex
Work on the introduction

The key novelty of our framework lies in its practical applicability. In practice, a SMR algorithm must support fast reconfiguration to replace faulty components, without which the long-term stability of a system is inevitably compromised. Our framework clearly separates the execution of commands from the reconfiguration decisions (including changing SMR module), allowing the use of an external reconfiguration master. This enables tolerating $k$ failures with $k+1$ replicas, a very
important improvement over the traditional $2k+1$ needed replicas.
Moreover, in practice, commands sent to a SMR algorithm often commute. Our framework is based on Generalized Consensus, a formulation of the State-Machine Replication problem that allows processing commutative commands without synchronization.

TODO: expand on the preceding paragraph.

Our framework is fully formalized and its composition theorem proved in Isabelle/HOL. Moreover, to quickly check concrete modules, we also provide a TLA+ specification of our framework which allows model-checking algorithms with the TLC model-checker.

TODO: the two abstract specs.

>>>>

The key novelty of our framework lies in its practical applicability, allowing the execution of commutative commands without synchronization and reducing the replication overhead from $2*k+1$ replicas to $k+1$ replicas, to tolerate $k$ replica crashes, using a reliable reconfiguration master.

Allowing the execution of commuting commands to proceed without synchronization is an important practical optimization, as commands may often commute in practice, for example when operating on disjoint parts of a data structure. 
To enable the synchronization-free execution of commuting commands, we have based our framework on Generalized Consensus \cite{Lamport05GeneralizeConsensus}. Traditional SMR implementations are implemented by having all the replicas execute the same sequence of commands, using a separate instance of a consensus algorithm to agree on the command to execute at each index of the command sequence. Using separate instances of a consensus algorithm for each index in the sequence of commands
makes it hard to optimize the execution of commuting command because commutativity is not a property of a single command.
Instead of restricting agreement on a single value, as in the consensus problem, in the  Generalized Consensus problem the set of replica servers must agree on the state of a growing data structure called a command-structure set. A command-structure set is a data structure with an append operation such that the order in which two commuting commands are appended does not change the resulting state. Thanks to this property, two replicas executing commuting commands in a different order will
not end up in a different state and can continue executing without further synchronization. This property can be leveraged to reduce the latency of processing one command to two message delays instead of three in message passing systems. As we will show, it also enables the processing of read-only commands in two message delays, although read-only commands do not commute with write commands.

In practice, a SMR algorithm must support fast reconfiguration to replace faulty components, without which the long-term stability of a system is inevitably compromised. Our framework clearly separates the execution of commands from the reconfiguration decisions (including changing SMR module), allowing the use of an external reconfiguration master. This enables tolerating $k$ failures with $k+1$ replicas, an important improvement over the traditional $2k+1$ needed replicas.

To demonstrate the practical applicability of our framework we provide composable versions of the generalized variations of popular algorithms like the classic Paxos, Fast Paxos, and multi-coordinated Paxos. We have checked the composability of these algorithms with the TLC model-checker and we provide all the specifications needed to replicate these experiments.

Our framework is also fully formalized and its composition theorem proved in the foundational interactive theorem prover Isabelle/HOL, certifying the correctness of the composition theorem at the deepest level.
